Bayesian Detection of Bias in Peremptory Challenges Using Historical Strike Data

### 2023-02-23 analysis/memo_TAS_reviewers.Rmd
small edit

4. Absent some very strong assumptions, there is no generally justifiable hyper prior for $\alpha$ in this context of strike bias and _Batson_ challenges. This is because the proper weight to place on historical data depends on what causes incompatibility between the historical data and the current data. The plausible reasons for non-random missingness and such confounding, however, likely vary considerably depending on strike procedure, courtroom practices, prosecutor charging practices, among other things. As we now note in the Discussion section of the paper, the literature on the power prior discusses information criteria for guide values of $\alpha$ as well as treating $\alpha$ as random and assigning a prior to it. However, that literature is unclear as to what implies for _Batson_ and peremptory strikes, for at least three reasons.  First, that literature has mostly focused on clinical trials and models whose properties are well known (e.g., the linear model), as distinct from the models of strike procedure discussed here. In the clinical trial setting, it might also be easier to capture(measure?) trial-level (and patient-level) characteristics that may be responsible for incompatibility. Even in these settings, selecting a prior for $\alpha$ is not straightforward. @Neelon2010 warn that using random power parameters (in their case, with a beta prior distribution) tends to attenuate the influence of the historical data, and so one might have to use highly informative priors on the power parameter. These authors suggest that practitioners consider whether "it is more appropriate to assign $\alpha$ a fixed value based on expert opinion about the relevance of the historical data to the current analysis". More recent versions of the power prior are beginning to address this issue, for example using "commensurate power priors" (@Hobbs2011).   Second, our approach does not attempt to maximize coverage in the traditional sense, but aims only to "detect" bias by determining whether zero falls outside the chosen credible interval for the posterior distribution of the bias parameter. Third, whereas some papers treat incompatibility as itself a parameter [e.g., @Ollier2020], the precise approaches rely on strong assumptions that do not obviously make sense in our context.  Accordingly, we believe the most conservative choice (for now) is to keep $\alpha$ a fixed value and check for sensitivity of results to the choice of $\alpha$ value. Proceeding otherwise would require the work of a completely new paper.

>>>>

4. Absent some very strong assumptions, there is no generally justifiable hyper prior for $\alpha$ in this context of strike bias and _Batson_ challenges. This is because the proper weight to place on historical data depends on what causes incompatibility between the historical data and the current data. The plausible reasons for non-random missingness and such confounding, however, likely vary considerably depending on strike procedure, courtroom practices, prosecutor charging practices, among other things. As we now note in the Discussion section of the paper, the literature on the power prior discusses information criteria for guide values of $\alpha$ as well as treating $\alpha$ as random and assigning a prior to it. However, that literature is unclear as to what implies for _Batson_ and peremptory strikes, for at least three reasons.  First, that literature has mostly focused on clinical trials and models whose properties are well known (e.g., the linear model), as distinct from the models of strike procedure discussed here. In the clinical trial setting, it might also be easier to capture(measure?) trial-level (and patient-level) characteristics that may be responsible for incompatibility. Even in these settings, selecting a prior for $\alpha$ is not straightforward. @Neelon2010 warn that using random power parameters (in their case, with a beta prior distribution) tends to attenuate (more downweighting) the influence of the historical data, and so one might have to use highly informative priors on the power parameter. These authors suggest that practitioners consider whether "it is more appropriate to assign $\alpha$ a fixed value based on expert opinion about the relevance of the historical data to the current analysis". More recent versions of the power prior are beginning to address this issue, for example using "commensurate power priors" (@Hobbs2011).   Second, our approach does not attempt to maximize coverage in the traditional sense, but aims only to "detect" bias by determining whether zero falls outside the chosen credible interval for the posterior distribution of the bias parameter. Third, whereas some papers treat incompatibility as itself a parameter [e.g., @Ollier2020], the precise approaches rely on strong assumptions that do not obviously make sense in our context.  Accordingly, we believe the most conservative choice (for now) is to keep $\alpha$ a fixed value and check for sensitivity of results to the choice of $\alpha$ value. Proceeding otherwise would require the work of a completely new paper.
